---
title: "DevVarShh"
author: "Luke Hayden"
date: "July 30, 2016"
output:
  pdf_document: default
  html_document: default
---

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
library("ggplot2")
library("reshape2")
library("ade4") 
library("ecodist")
library("phytools")

setwd("~/varpap")
ed <- read.csv("ed.csv")
setwd("~/varpap/mx")


ed$nid <- paste(ed$Weight, ed$Litter, sep="")


```


#Import
```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

mxstg <- read.table("mxFVBDUHiCD1scores20.txt")



mxstg$mxstgnid <- paste(mxstg$Weight, mxstg$Clutch, sep="")

add <- subset(mxstg,  !(mxstg$mxstgnid %in% ed$nid) & mxstg$Strain != "CD1")




ed <- ed[order(ed$nid),]

ed$aw <- paste(ed$Age.dpc., ed$Weight)

add$aw <- paste(add$Age, add$Weight)

write.csv(add, file="add.csv")


ged <- subset(ed, ed$Litter %in% mxstg$Clutch)
ged <- subset(ged, ged$nid %in% mxstg$mxstgnid)



mxstg <- mxstg[order(mxstg$mxstgnid),]
mxstgdf <- subset(mxstg, mxstg$Strain == "DUHi" | mxstg$Strain == "FVB" )
mxstgdf <- subset(mxstgdf, mxstgdf$mxstgnid %in% ged$nid)
mxstgdf <- cbind(mxstgdf, ged[,c(2,9:12)])
mxstgdf <- subset(mxstgdf, mxstgdf$Age == mxstgdf$Age.dpc.)

mxstgmelted <- melt(mxstg, id.vars=c("ID", "Strain", "Age", "Weight", "Clutch", "shh_M2_spot", "M1_spot_split_LR", "Total_state", "stateord", "Symmetry", "groupwt"))
  
mxstgdfmelted <- melt(mxstgdf, id.vars=c("ID", "Strain", "Age", "Weight", "Clutch", "shh_M2_spot", "M1_spot_split_LR", "Total_state", "stateord", "Symmetry", "groupwt", "mxstgnid", "nid", "mediantemps", "bornemin", "bornemax"))

mxstgdfmelted$variable[mxstgdfmelted$variable == "shh_R2_spot"] <- "R2 Shh expression"
mxstgdfmelted$variable[mxstgdfmelted$variable == "shh_M1_spot"] <- "M1 Shh expression"
mxstgdfmelted$variable[mxstgdfmelted$variable == "cap_transition"] <- "Cap transition"
mxstgdfmelted$variable[mxstgdfmelted$variable == "shboob"] <- "Anterior protrusion"

```


#Logistic regression
```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

#logistic regression
subsmxstg <- subset(mxstg, Strain!= "CD1")
subsmxstg$shh_R2_spot <- factor(subsmxstg$shh_R2_spot > 0)
subsmxstg$Strain <- factor(subsmxstg$Strain)
mxlogitR2 <- glm(shh_R2_spot ~ Weight + Strain, data = subsmxstg, family = "binomial")

subsmxstg$shh_M1_spot <- factor(subsmxstg$shh_M1_spot > 0)
mxlogitM1 <- glm(shh_M1_spot ~ Weight + Strain, data = subsmxstg, family = "binomial")
summary(mxlogitM1)

subsmxstg$cap_transition <- factor(subsmxstg$cap_transition > 0)
mxlogitCAP <- glm(cap_transition ~ Weight + Strain, data = subsmxstg, family = "binomial")
summary(mxlogitCAP)

mxlogitSHB <- glm(shboob ~ Weight + Strain, data = subsmxstg, family = "binomial")
summary(mxlogitSHB)

```

#Graphing

#Each character in its own plot
```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
ggplot(mxstgmelted, aes(x=Weight,y=value,color=Strain,shape=Strain)) +geom_point(size=1.5) + facet_wrap(~ variable,ncol=1) +  scale_y_continuous(breaks=c(0,1,2)) + scale_colour_manual(values=c("forest green", "red3", "cornflowerblue")) +xlab("Embryo Weight(mg)")+ylab("")
dev.copy(svg,'charplot.svg')
dev.off()
dev.copy(pdf,'charplot.pdf')
dev.off()
dev.copy(png,'charplot.png')
dev.off()

```

with calculated devtime
```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
  
mxstg <- mxstg[order(mxstg$mxstgnid),]
mxstgdf <- subset(mxstg, mxstg$Strain == "DUHi" | mxstg$Strain == "FVB" )
mxstgdf <- subset(mxstgdf, mxstgdf$mxstgnid %in% ged$nid)
mxstgdf <- cbind(mxstgdf, ged[,9:12])

colnames(mxstgdf) <-  c("ID","Strain"    ,       "Age"      ,        "Weight"    ,       "Clutch"          
 ,"R2 Shh expression"   ,   "M1 Shh expression"    ,  "shh_M2_spot"  ,    "Cap transition" ,  "M1_spot_split_LR"
, "Anterior protrusion"     ,      "Total_state"   ,   "stateord"    ,     "Symmetry"    ,     "groupwt"         
, "mxstgnid"      ,   "mediantemps"   ,   "bornemin"    ,     "bornemax"  ,       "nid")
mxstgdfmelted <- melt(mxstgdf, id.vars=c("ID", "Strain", "Age", "Weight", "Clutch", "shh_M2_spot", "M1_spot_split_LR", "Total_state", "stateord", "Symmetry", "groupwt", "mxstgnid", "nid", "mediantemps", "bornemin", "bornemax"))

mxstgdfmelted$variable[mxstgdfmelted$variable == "shh_R2_spot"] <- "R2 Shh expression"
mxstgdfmelted$variable[mxstgdfmelted$variable == "shh_M1_spot"] <- "M1 Shh expression"
mxstgdfmelted$variable[mxstgdfmelted$variable == "cap_transition"] <- "Cap transition"
mxstgdfmelted$variable[mxstgdfmelted$variable == "shboob"] <- "Anterior protrusion"

mxstgdfmelted$value[mxstgdfmelted$Strain == "FVB"] <- mxstgdfmelted$value[mxstgdfmelted$Strain == "FVB"] +0.05
mxstgdfmelted$value[mxstgdfmelted$Strain == "DUHi"] <- mxstgdfmelted$value[mxstgdfmelted$Strain == "DUHi"] -0.05


ggplot(mxstgdfmelted, aes(x=mediantemps,y=value,color=Strain,shape=Strain)) +geom_point(size=1.5) + facet_wrap(~ variable,ncol=1) +  scale_y_continuous(breaks=c(0,1,2)) + scale_colour_manual(values=c("red3",  "cornflowerblue")) +xlab("Computed developmental age (cdpc)")+ylab("") +theme_bw()

ggsave(plot=p, height=10, width=7, file="borneidentity.pdf", device=pdf)




```

no cd1, DUHi and FVB only

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

nocdmxstgmelted <- subset(mxstgmelted, Strain != "CD1")
ggplot(nocdmxstgmelted, aes(x=Weight,y=value,color=Strain,shape=Strain)) +geom_point(size=1.5) + facet_wrap(~ variable,ncol=1) +  scale_y_continuous(breaks=c(0,1,2)) + scale_colour_manual(values=c("red3", "cornflowerblue"))

ggplot(nocdmxstgmelted, aes(x=Weight,y=value,color=Strain,shape=Strain)) +geom_point(size=1.5) + facet_wrap(~ variable,ncol=1) +  scale_y_continuous(breaks=c(0,1,2)) + scale_colour_manual(values=c("forest green", "cornflowerblue"))
dev.copy(svg,'nocdcharplot.svg')
dev.off()
dev.copy(pdf,'nocdcharplot.pdf')
dev.off()
dev.copy(png,'nocdcharplot.png')
dev.off()

```

#Total state, all strains together

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
mxtot <- melt(mxstg, id.vars=c("ID", "Strain", "Age", "Weight", "Clutch", "shh_M2_spot", "M1_spot_split_LR", "shh_R2_spot", "shh_M1_spot", "shboob", "cap_transition", "Total_state", "Symmetry", "groupwt")) 
ggplot(mxtot, aes(x=Weight,y=value,color=Strain,shape=Strain)) +geom_point(size=2) + facet_wrap(~ variable,ncol=1) + scale_colour_manual(values=c("forest green", "red3", "cornflowerblue")) +xlab("Embryo Weight(mg)")+ylab("")

dev.copy(svg,'totstateallstr.svg')
dev.off()

```

#Total state, one panel by strain

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

mxstgpanbystr <- melt(mxstg, id.vars=c("ID", "Strain", "Age", "Weight", "Clutch", "shh_M2_spot", "M1_spot_split_LR", "Total_state", "shh_R2_spot", "shh_M1_spot", "shboob", "cap_transition", "Symmetry", "groupwt")) 


ggplot(mxstgpanbystr, aes(x=Weight,y=value,color=Strain)) +geom_point(size=1) + facet_wrap(~ Strain,ncol=1) + scale_colour_manual(values=c("forest green", "red3", "cornflowerblue")) +xlab("Embryo Weight(mg)") +ylab("")
dev.copy(svg,'totstatepanbystr.svg')
dev.off()
dev.copy(pdf,'totstatepanbystr.pdf')
dev.off()


mxstgpanbystr <- melt(mxstg, id.vars=c("ID", "Strain", "Age", "Weight", "Clutch", "shh_M2_spot", "M1_spot_split_LR", "Total_state", "shh_R2_spot", "shh_M1_spot", "shboob", "cap_transition", "Symmetry", "groupwt")) 

mxstgdfmelted <- melt(mxstgdf, id.vars=c("ID", "Strain", "Age", "Weight", "Clutch", "shh_M2_spot", "M1_spot_split_LR", "Total_state", "stateord", "Symmetry", "groupwt", "mxstgnid", "nid", "mediantemps", "bornemin", "bornemax"))
````

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}


m2 <- data.frame(id =mxstgdf$ID, totstat=mxstgdf$Total_state,Strain=mxstgdf$Strain, stateord =mxstgdf$stateord, weight=mxstgdf$Weight, cdpc=mxstgdf$mediantemps)
m2 <- subset(m2, m2$Strain != "CD1")
d <- subset(m2, m2$Strain =="DUHi" )
y=levels(m2$stateord)
f <- subset(m2, m2$Strain =="FVB")

i <- d$stateord %in% f$stateord
d2 <- subset(d, !i)


i <- f$stateord %in% d$stateord
f2 <- subset(f, !i)

m3 <- data.frame(
  y=levels(m2$stateord), 
  din =y %in% d2$stateord,
  fin=y %in% f2$stateord,
  end=max(m2$cdpc),
  start=min(m2$cdpc)) 


m4 <- melt(m3, id.vars=c("y", "end", "start"))
m4$Strain <- "DUHi"
m4$Strain[m4$variable == "fin"] <- "FVB"

ggplot(m2, aes(x=cdpc,y=stateord,color=Strain)) +
  geom_point(size=1) + 
  facet_wrap(~ Strain,ncol=1) + 
  scale_colour_manual(values=c("red3",  "red3", "cornflowerblue","cornflowerblue"))   +
  xlab("Computed Embryo Age (cdpc)") +
  ylab("Embryo overall state") +
  geom_segment(inherit.aes=FALSE, data=m4, aes(x=start, xend=end, y=y, yend=y, alpha=value, colour=variable, size=10)) +
  scale_alpha_manual(values=c(0,0.3)) +
  theme_bw()+
  theme(legend.position = "none")


p=ggplot(m2, aes(x=cdpc,y=stateord,color=Strain)) +
  geom_point(size=1) + 
  facet_wrap(~ Strain,ncol=1) + 
  scale_colour_manual(values=c("red3",  "red3", "cornflowerblue","cornflowerblue"))   +
  xlab("Computed Embryo Age (cdpc)") +
  ylab("Embryo overall state") +
  geom_segment(inherit.aes=FALSE, data=m4, aes(x=start, xend=end, y=y, yend=y, alpha=value, colour=variable, size=10)) +
  scale_alpha_manual(values=c(0,0.3)) +
  theme_bw()+
  theme(legend.position = "none")

ggsave(plot=p, filename="ovstate.png", device="png", height=6, width=6)

ggsave(plot=p, filename="ovstate.svg", device="svg", height=6, width=6)


```



```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

ggplot(mxstgdfmelted, aes(x=Age, y=mediantemps)) +geom_point()
````

#Weight and dev state dist matrices for DUHI
```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
mxsd <- subset(mxstg, Strain=="DUHi")


weights <- (mxsd[,4])
R2 <- (mxsd[,6])
M1 <- (mxsd[,7])
CAP <- (mxsd[,8])
SHB <- (mxsd[,9])
IDs <- (mxsd[,1])

wtdmat <- dist(weights)
r2dmat <- dist(R2)
mr2dmat <- as.matrix(r2dmat)
m1dmat <- dist(M1)
mm1dmat <- as.matrix(m1dmat)
capdmat <- dist(CAP)
mcapdmat <- as.matrix(capdmat)
shbdmat <- dist(SHB)
mshbdmat <- as.matrix(shbdmat)
statedmat <- r2dmat + m1dmat + capdmat + shbdmat
vstatedmat <- c(statedmat)
mstatedmat <- as.matrix(statedmat)

vwtdmat <- c(wtdmat)
isclosemat <- wtdmat <20

dstatedmat <- statedmat
dwtdmat <- wtdmat
dvstatedmat <- vstatedmat
dvwtdmat <- vwtdmat
disclosemat <- isclosemat

mwtdmat <- as.matrix(wtdmat)


#######Timeline of dev differences -matrix for DUHi
closedatfr <- data.frame(1,1,0,0,0,0,0,0)
closedatfr
x <- 1
y <- 1
p1wt <- 0
statediff <- 0
#colnames(closedatfr) <- c("p1", "p2", "p1wt", "statediff")
#closedatfr[1,1]
subindex <- c(1:(length(weights)))
#closedatfr <- rbind(closedatfr, c(y, x, p1wt, statediff))

while (y < length(weights)) {
  p1wt <- mxsd[y,4]
  neighbourWTs <- weights[mwtdmat[y,] < 10]
  neighbourIndices <- subindex[mwtdmat[y,] < 10]
  x <- 1
  while (x < length(neighbourWTs)) {
    xindex <- neighbourIndices[x]
    closedatfr <- rbind(closedatfr, c(y, xindex, p1wt, mstatedmat[y,xindex], mr2dmat[y,xindex], mm1dmat[y,xindex], mcapdmat[y,xindex], mshbdmat[y,xindex]))
    x <- x+1
  }  
  y <- y+1
}

closedatfr[,9] <- "DUHi"

#closedatfr <- cbind(closedatfr, "DUHi")
colnames(closedatfr) <- c("p1", "p2", "p1wt", "statediff", "R2diff", "M1diff", "capdist", "shbdist", "Strain")
#closedatfr[100:,5] <- "FVB"


#######

closedatfr <- closedatfr[-c(1),]
duclosedatfr <- closedatfr


```


#Weight and dev state dist matrices for FVB
```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

mxsd <- subset(mxstg, Strain=="FVB")

####Weight and dev state dist matrices for FVB
weights <- (mxsd[,4])
R2 <- (mxsd[,6])
M1 <- (mxsd[,7])
CAP <- (mxsd[,8])
SHB <- (mxsd[,9])
IDs <- (mxsd[,1])

wtdmat <- dist(weights)
r2dmat <- dist(R2)
mr2dmat <- as.matrix(r2dmat)
m1dmat <- dist(M1)
mm1dmat <- as.matrix(m1dmat)
capdmat <- dist(CAP)
mcapdmat <- as.matrix(capdmat)
shbdmat <- dist(SHB)
mshbdmat <- as.matrix(shbdmat)
statedmat <- r2dmat + m1dmat + capdmat + shbdmat
vstatedmat <- c(statedmat)
mstatedmat <- as.matrix(statedmat)
vwtdmat <- c(wtdmat)
mwtdmat <- as.matrix(wtdmat)


fstatedmat <- statedmat
fwtdmat <- wtdmat
fvstatedmat <- vstatedmat
fvwtdmat <- vwtdmat
fisclosemat <- isclosemat


#######Timeline of dev differences -matrix for FVB
closedatfr <- data.frame(1,1,0,0,0,0,0,0)
#closedatfr
x <- 1
y <- 1
p1wt <- 0
statediff <- 0
#colnames(closedatfr) <- c("p1", "p2", "p1wt", "statediff")
#closedatfr[1,1]
subindex <- c(1:(length(weights)))
#closedatfr <- rbind(closedatfr, c(y, x, p1wt, statediff))
xindex <- 1

while (y < length(weights)) {
  p1wt <- mxsd[y,4]
  neighbourWTs <- weights[mwtdmat[y,] < 10]
  neighbourIndices <- subindex[mwtdmat[y,] < 10]
  x <- 1
  while (x < length(neighbourWTs)) {
    xindex <- neighbourIndices[x]
    closedatfr <- rbind(closedatfr, c(y, xindex, p1wt, mstatedmat[y,xindex], mr2dmat[y,xindex], mm1dmat[y,xindex], mcapdmat[y,xindex], mshbdmat[y,xindex]))
    x <- x+1
  }  
  y <- y+1
}

closedatfr[,9] <- "FVB"

#closedatfr <- cbind(closedatfr, "DUHi")
colnames(closedatfr) <- c("p1", "p2", "p1wt", "statediff", "R2diff", "M1diff", "capdist", "shbdist", "Strain")
#closedatfr[100:,5] <- "FVB"

```
#Plotting developmental varability matrix over time

With binning
```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

#bothclosedatfr[,10] <- (as.integer((bothclosedatfr$p1wt/(max(bothclosedatfr$p1wt)-min(bothclosedatfr$p1wt)))*20))*5

fvclosedatfr <- closedatfr
bothclosedatfr <- rbind(duclosedatfr,fvclosedatfr)
bothclosedatfr[,10] <- as.integer(bothclosedatfr$p1wt/10)*10
upbins <- (as.integer(bothclosedatfr$p1wt/10)*10)+10
bothclosedatfrplus <- bothclosedatfr
bothclosedatfrplus$V10 <- upbins
#bothclosedatfrplus <- rbind(bothclosedatfr, bothclosedatfrclone)

colnames(bothclosedatfrplus) <- c("p1", "p2", "p1wt", "statediff", "R2diff", "M1diff", "capdist", "shbdist", "Strain", "bin")


a <- min(bothclosedatfrplus$bin)
bothgrps <- c(0)
dubins <- c()
fvbins <- c()
binlist <- c()
 
while (a < max(bothclosedatfrplus$bin)) {
  s <- subset(bothclosedatfrplus, bin==a)
  td <- subset(s, Strain == "DUHi")
  dumea <- 0+ mean(td$statediff)
  dubins <- c(dubins, dumea)
  tf <- subset(s, Strain == "FVB")
  fvmea <- 0+ mean(tf$statediff)
  fvbins <- c(fvbins, fvmea)
  binlist <- c(binlist, a)
  a <- a +10 
}

fvbins


plot(binlist, fvbins, col="cornflowerblue", type="o", pch=16, bg="cornflowerblue", lty=1, xlim = c(80,350), ylim = c(0,4))
lines(binlist, dubins, type="o", col="green", pch=1, lty=1)
lines(duclosedatfr$p1wt, duclosedatfr$statediff, type="o", col="red3", pch=0, lty=0 )
lines(fvclosedatfr$p1wt, fvclosedatfr$statediff, type="o", col="cornflowerblue", pch=0, lty=0 )

dudat <- data.frame(dubins, binlist, c("DUHi"))
colnames(dudat) <- c("meandiff", "wtbin", "strain")
fvdat <- data.frame(fvbins, binlist, c("FVB"))
colnames(fvdat) <- c("meandiff", "wtbin", "strain")
bothdat <- rbind(dudat, fvdat)

ggplot(bothdat, aes(x=wtbin, y=meandiff, color=strain)) + geom_line(size=2) +xlim(90,400) +ylim(0,2) + scale_colour_manual(values=c("red3", "cornflowerblue")) +ylab("Mean Developmental State Difference") +xlab("Embryo Weight(mg)") + coord_fixed(ratio=35)
dev.copy(svg,'devdisttimelineslidewindow.svg')
dev.off()
dev.copy(pdf,'devdisttimelineslidewindow.pdf')
dev.off()


```


with loess
```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}


####Plotting developmental variation over time w/ loess

ggplot(bothclosedatfr, aes(x=p1wt,y=statediff, color=Strain)) +geom_point(size=2) + stat_smooth(size=1.5) +xlim(90,400) + scale_colour_manual(values=c("red3", "cornflowerblue")) +ylab("Mean Developmental State Difference") +xlab("Embryo Weight(mg)")
ggplot(bothclosedatfr, aes(x=p1wt,y=statediff, color=Strain)) +geom_point(size=0, alpha =0) + stat_smooth(method=loess, size=1.5) +xlim(90,380) + ylim(-0.5,2)+ scale_colour_manual(values=c("red3", "cornflowerblue")) +ylab("Mean Developmental State Difference") +xlab("Embryo Weight(mg)") + coord_fixed(ratio=40)
dev.copy(svg,'devdisttimelineloess.svg')
dev.off()
dev.copy(pdf,'devdisttimelineloess.pdf')
dev.off()
dev.copy(png,'devdisttimelineloess.png')
dev.off()

mbothclosedatfr <- melt(bothclosedatfr, id.vars=c("p1",  "p2",  "p1wt", "Strain", "V10"))
ggplot(mbothclosedatfr, aes(x=p1wt,y=value,color=Strain)) +geom_point(size=0) + facet_wrap(~ variable,ncol=1) + stat_smooth(method=loess, size=1.5)
dev.copy(svg,'devdisttimelinebychar.svg')
dev.off()
dev.copy(pdf,'devdisttimelinebychar.pdf')
dev.off()

```


#Plotting state difference versus weight difference
```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

########
plot(dvwtdmat, dvstatedmat, lty=0, pch="I", cex=4,col=adjustcolor("forest green", alpha=0.05), xlab="Age difference(mg)", ylab="Developmental state difference", main="DUHi")
dev.copy(svg,'DUdevdistVweightdist.svg')
dev.off()
dev.copy(pdf,'DUdevdistVweightdist.pdf')
dev.off()

dev.copy(png,'DUdevdistVweightdistg.png')
dev.off()


plot(fvwtdmat, fvstatedmat, lty=0, col=adjustcolor("cornflowerblue", alpha=0.05), pch="I", cex=4, xlab="Age difference(mg)", ylab="Developmental state difference", main="FVB")
dev.copy(svg,'FVdevdistVweightdist.svg')
dev.off()
dev.copy(pdf,'FVdevdistVweightdist.pdf')
dev.off()
dev.copy(png,'FVdevdistVweightdist.png')
dev.off()



plot(fvwtdmat, fvstatedmat, lty=0, col=adjustcolor("blue", alpha=0.005), pch=15, cex=4, xlab="Age difference(mg)", ylab="Developmental state difference", xlim =c(0,400))
lines(dvwtdmat, dvstatedmat, type="p", pch=15, cex=4,col=adjustcolor("red", alpha=0.006))
lines(fvwtdmat, fvstatedmat, type="p", pch=15, cex=4,col=adjustcolor("blue", alpha=0.005))
#lines(dvwtdmat, dvstatedmat, type="p", pch=15, cex=4,col=adjustcolor("red3", alpha=0.01))
#lines(fvwtdmat, fvstatedmat, type="p", pch=15, cex=4,col=adjustcolor("cornflower blue", alpha=0.005))

legend(320,1.5,  c("DUHi","FVB"), lty=c(1,1), lwd=c(2,2),col=c("red","cornflowerblue"))
dev.copy(svg,'bothdevdistVweightdist.svg')
dev.off()
dev.copy(pdf,'bothdevdistVweightdist.pdf')
dev.off()

```


#Mantel
```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

mxfmant <- mantel(formula = fwtdmat ~ fstatedmat, data = sys.parent(), nperm = 1000,
       mrank = FALSE, nboot = 500, pboot = 0.9, cboot = 0.95)
summary(mxfmant)
mxfmant
mxdmant <- mantel(formula = dwtdmat ~ dstatedmat, data = sys.parent(), nperm = 1000,
                mrank = FALSE, nboot = 500, pboot = 0.9, cboot = 0.95)
summary(mxdmant)
mxdmant

dpmant <- multi.mantel(dwtdmat, dstatedmat, nperm=1000)
dpmant
fpmant <- multi.mantel(fwtdmat, fstatedmat, nperm=1000)
fpmant
```

Plotting residuals from the Mantel

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

plot(dpmant$residuals, col="red3", type="o", pch=3, ylab="Residuals", main="DUHi")
dev.copy(svg,'dudevdistVweightdistresidplot.svg')
dev.off()
dev.copy(pdf,'dudevdistVweightdistresidplot.pdf')
dev.off()


plot(fpmant$residuals, col="cornflowerblue", type="o", pch=3, lty=0, ylab="Residuals", main="FVB")
dev.copy(svg,'fvdevdistVweightdistresidplot.svg')
dev.off()
dev.copy(pdf,'fvdevdistVweightdistresidplot.pdf')
dev.off()
```

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}


dreds <- c(dpmant$residuals)

freds <- c(fpmant$residuals)
```

 

#Wilcoxon rank-sum test on the residuals
```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
wilcox.test(dreds[dreds >0], freds[freds >0])
wilcox.test(abs(freds), abs(dreds))

```

#Boxplot of total developmental variation 
Takes the positive residuals from each Mantel test of the developmental state distance matrix vs the embryo weight distance matrix

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
ddu <- data.frame(c("DUHi"), dreds[dreds >0])
colnames(ddu) <- c("Strain", "resid")
ffu <- data.frame(c("FVB"), freds[freds >0])
colnames(ffu) <- c("Strain", "resid")

resids <- rbind(ddu,ffu)


qplot(Strain, resid, data = resids, geom = "boxplot",color=Strain) + scale_colour_manual(values=c("red3", "cornflowerblue"))+ylab("Developmental Variation")
dev.copy(svg,'devvarboxplotplot.svg')
dev.off()
dev.copy(pdf,'devvarboxplotplot.pdf')
dev.off()

```

